# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
data <- read.csv("activity/activity.csv")
```

## Histogram of the total number of steps taken each day
```{r}
stepsPerDay =  tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(stepsPerDay, xlab="steps per day")
title(xlab="steps per day")
```
## Mean total number of steps taken per day 
```{r}
mean(stepsPerDay)

```

## Median total number of steps taken per day
```{r}
median(stepsPerDay)
```


## Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
avgNumStepsPerInterval <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
plot(names(avgNumStepsPerInterval), avgNumStepsPerInterval, type="l", xlab="5-minute Intervals", ylab="Average Number of Steps Taken")
```

## 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps
```{r}
names(avgNumStepsPerInterval)[which.max(avgNumStepsPerInterval)[[1]]]
```


## Imputing missing values

##  total number of missing values in the dataset
```{r}
sum(is.na(data$steps))
```

## strategy for filling in all of the missing values in the dataset
### missing values in steps column are filled up with mean number of steps of corresponding 5-minute intervals.

##  new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
nas <- which(is.na(data$steps))
avgNumStepsPerIntervalFrame <- as.data.frame(as.table(avgNumStepsPerInterval))
avgNumStepsPerIntervalFrame[,1] = as.numeric(as.character(avgNumStepsPerIntervalFrame[,1]))
data$steps[nas] = sapply(nas, function(x) 
                    avgNumStepsPerIntervalFrame[,2][avgNumStepsPerIntervalFrame[,1]==data$interval[x]])

head(data)
```

## Histogram of the total number of steps taken each day with NA's filled in 
```{r}
stepsPerDay =  tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(stepsPerDay, xlab="steps per day")
title(xlab="steps per day")
```
## Mean total number of steps taken per day with NA's filled in
```{r}
mean(stepsPerDay)

```

## Median total number of steps taken per day with NA's filled in
```{r}
median(stepsPerDay)
```

## Impact of imputing missing data on the estimates of the total daily number of steps?
### imputing missing data increase the average total daily number of steps as expected



## Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
bool <- (weekdays(as.Date(data$date))=="Sunday") | ((weekdays(as.Date(data$date))=="Saturday"))
data$day = "Weekday"
data$day[which(bool)] = "Weekend"
head(data)
```

## Making a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
avg1 <- tapply(data$steps[which(data$day=="Weekday")], data$interval[which(data$day=="Weekday")], mean)
avg2 <- tapply(data$steps[which(data$day=="Weekend")], data$interval[which(data$day=="Weekend")], mean)
avg1 <- as.data.frame(as.table(avg1))
avg2 <- as.data.frame(as.table(avg2))
avg1$day="Weekday"
avg2$day="Weekend"
avg <- rbind(avg1, avg2)
head(avg)
avg$Var1 <- as.numeric(as.character(avg$Var1))
library(lattice)
xyplot(Freq ~ Var1 | day, data=avg, layout=c(1,2), type='l')
```





